---
title: "Practical Machine Learning Course Project"
author: "hcam"
date: "2/9/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practical Machine Leraning Course Project

In this assignment we will explore the accelerometer data from the belt, forearm, arm, and dumbell of 6 participants. These participants were asked to perform exercises correctly and incorrectly in 5 different ways. We will use this data to build a model to predict if a user was doing an exercise correctly or not.

Data used for this assignment is courtesy of http://groupware.les.inf.puc-rio.br/har. 

## Loading The Required Packages

```{r packages, results="hide", message=FALSE, warning = FALSE}
library(tidyverse)
library(caret)
library(randomForest)
library(rpart)
library(rattle)
```

## Loading the Data

The training and testing data sets are available the links below.  I will utilize the tidyverse package to load the data via read_csv function.

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r data, results="hide", message=FALSE, warning=FALSE}
train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

train <- read_csv(train_url)
test <- read_csv(test_url)
```

```{r dim_raw}
dim(train)
dim(test)
```
##Preparing the Data

We will select the columns without NA values, and also we will filter out the first seven columns of data.  

```{r prep_data, message=FALSE}
train <- train[, colSums(is.na(train)) == 0] #subset the data for non NA values
test <- test[, colSums(is.na(test)) == 0] 

train <- train[,-(1:7)] #removing the first seven columns of data.
test <- test[,-(1:7)]
```

```{r dim}
dim(train)
dim(test)
```

###Partioning the Training Data
```{r, part_data, message = FALSE}
set.seed(555)
inTrain <- createDataPartition(train$classe, p = 0.75, list = FALSE)
training <- train[inTrain,]
testing <- train[-inTrain,]
```

```{r part_dim}
dim(training)
dim(testing)
```

We will also remove the Zero Covariates. These variables have low variablity and are not useful in our model.

```{r nearzero, message = FALSE }
nzv <- nearZeroVar(training, saveMetrics = TRUE)
training <- training[,nzv$nzv == FALSE]
testing <- testing[,nzv$nzv == FALSE]
```

###Building our Model

We will use the Random Forest method for our model.  This is one of the most used and accurate algorthims. 

```{r rf_model}
modFit <- train(classe ~ ., data=training, method="rf", list = FALSE, 
                trControl = trainControl(method="cv"), number = 5, allowParallel = TRUE)
```

Now we can test the model on our validation training set.

```{r validation}
testing_pred <- predict(modFit, testing)
confusionMatrix(testing$classe, testing_pred)
```

We can check the accuracy and the out of sample error rate below.

```{r accuracy, echo = FALSE}
as.numeric(confusionMatrix(testing$classe, testing_pred)$overall[1])
1-as.numeric(confusionMatrix(testing$classe, testing_pred)$overall[1])
```

###Prediction on Test Data

We can now apply our model on the test data set provided.
```{r final}
test_pred <- predict(modFit, test)
test_pred
```


